Cooperative Kernels: Safe Multitasking for Blocking Algorithms on GPUs

### 2016-11-15 write-ups/PLDI/paper.tex
Some work on experiments.

\subsection{Performance and Responsiveness of Cooperative Scheduling}\label{sec:responsiveness}

Cooperative scheduling enables to run a non-cooperative kernel alongside
a cooperative one, we focus here on the performance impact of this
multitasking and on the scheduling responsiveness. The performance is
measured with respect to the performance of each kernel run in
sequence. Responsiveness, i.e., how fast a non-cooperative kernel can be
scheduled, depends not only on the scheduler implementation, but also on
how often the cooperative kernel offer workgroups to be killed.

@@

\paragraph{Experimental setup}
For a given cooperative kernel and its data set, we make it run while
regularly asking to schedule a non-cooperative kernel that simulates
graphics rendering, according to the classes listed in
\mytab~\ref{tab:noncooperativeworkload}. In practice, we use a matrix
multiplication kernel as the non-cooperative workload, with matrix
dimensions tailored for each GPU to reach the appropriate execution
duration.  In separate runs, we ask the scheduler to launch this
non-cooperative kernels with different number of workgroups: either 1, a
quarter, a half or all but one workgroups among the ones available for
the cooperative kernel. Moreover, for graph algorithms we try both the
regular and the query barrier.

Our experiments span 13 pairs of cooperative kernels and data sets, 3
classes of non-cooperative kernel workloads and frequency, 4 quantities
of workgroups claimed for the non-cooperative kernel and 2 variations of
resizing barriers for graph algorithms, leading to 288
configurations. Each of this configuration is run 10 times on each
GPU. For each run, we record the execution time of the cooperative
kernel. For each scheduling of the non-cooperative kernel during the
run, we also record the \emph{gather time} needed by the scheduler to
collect enough workgroups to launch the non-cooperative kernel, and the
non-cooperative kernel execution time. We make sure all kernels perform
their computations by comparing their result with expected results. We
had to ignore some configurations where the non-cooperative kernel asks
for all but one workgroups (namely bc, bfs and l-sssp on all chips, and
color on Iris), and the game application on HD520, since they lead to
system issues like GPU deadlock or blue screens---which we tend to think
are manifestations of driver bugs.

@@

\paragraph{Results}  \ADComment{Discuss the results.}

\myfig~\ref{fig:fine-grained-timing} reports the measure for three
configurations and represents the gather and execution time of the
non-cooperative kernel w.r.t. the quantity of workgroups allocated to
it. Times are in milliseconds, and a logarithmic scale is used since
gather times tend to be much smaller than execution times. The
horizontal grey lines indicates the desired rate for non-cooperative
kernels.

>>>>

\subsection{Multitasking via Cooperative Scheduling}\label{sec:responsiveness}

We now assess the performance impact of multitasking between a
coopoerative and non-cooperative kernel, and on scheduling
responsiveness.  The performance is measured with respect to the
performance of each kernel run in sequence. Responsiveness, i.e., how
fast a non-cooperative kernel can be scheduled, depends not only on
the scheduler implementation, but also on how often the cooperative
kernel offer workgroups to be killed.

@@

\paragraph{Experimental setup}
For a given cooperative kernel and its data set, we launch the kernel
asynchronously and then repeatedly schedule a non-cooperative kernel
that aims to simulate the intensity of a graphics rendering workload,
according to the classes listed in
\mytab~\ref{tab:noncooperativeworkload}. In practice, we use a matrix
multiplication kernel as the non-cooperative workload, with matrix
dimensions tailored for each GPU to reach the appropriate execution
duration.  We conduct separate runs where we vary the number of workgroup slots requested by the non-cooperative kernel,
considering the cases where one, a quarter, a half, and all-but-one, of the total number of workgroup slots is requested.
Moreover, for the graph algorithms we try both the
regular and query barrier implementations.

Our experiments span 13 pairs of cooperative kernels and data sets, 3
classes of non-cooperative kernel workloads \ADComment{cut ``and frequency''?} and frequency, 4 quantities
of workgroups claimed for the non-cooperative kernel and 2 variations of
resizing barriers for graph algorithms, leading to 288
configurations.  We run each configuration 10 times on each
GPU, in order to report averaged performance numbers. For each run, we record the execution time of the cooperative
kernel. For each scheduling of the non-cooperative kernel during the
run, we also record the \emph{gather time} needed by the scheduler to
collect enough workgroups to launch the non-cooperative kernel, and the
non-cooperative kernel execution time. We sanity-check whether all kernels perform
their computations corretly by comparing their computed result with expected reference results.

To work around compiler crash and compiler hang bugs in Intel's current OpenCL drivers, we had to apply what semantics-preserving changes to the code generated by \kernelmerge{} for a few applications configurations.  We could not obtain results for some configurations where the non-cooperative kernel asks
for all but one workgroups (namely bc, bfs and l-sssp on all chips, and
color on Iris), and could not get the game application to run at all on HD520; these configurations led to 
catastrophic issues such as display freezes or blue screens; we also attribute these to bugs in Intel's current drivers.

@@

\paragraph{Results}

\myfiglong\ref{fig:fine-grained-timing} reports the measure for three
configurations and represents the gather and execution time of the
non-cooperative kernel w.r.t. the quantity of workgroups allocated to
it. Times are in milliseconds, and a logarithmic scale is used since
gather times tend to be much smaller than execution times. The
horizontal grey lines indicates the desired rate for non-cooperative
kernels.
